---
title: "Intro to Package Binomial"
author: "Lucy Brooke"
date: "5/5/2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro to Package Binomial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, echo = TRUE}
knitr::opts_chunk$set(echo = TRUE, error = TRUE, purl = FALSE)
library(binomial)
```

## Welcome to the package `"binomial"`!

`"binomial"` is a package that computes and visualizes different measures of given binomial distributions. 


## Function bin_choose()

This function allows the user to find the number of combinations in which given successes can occur in a fixed number of trials.

For example:
Say that you are flipping a fair coin. You want to know, given ten flips (n = 10), how many times is it possible to get tails six out of ten times (k = 6) ? Note: this function does not take into account the order of the successes. 

```{r}
bin_choose(n = 10, k = 6)
```

This means that there are 210 possible ways to get tails 6 times out of 10 flips.

## Function bin_probability()

This function allows the user to calculate the probability of success within a fixed number of trials. This is true as long as the probability of success remains the same for each iteration of the trial.

So -- to continue from the last example -- imagine that you now want to know the probability that you will get six tails in ten flips. Here, success = 6, trials = 10, and, because this is a fair coin, prob = 0.5.

```{r}
bin_probability(success = 6, trials = 10, prob = 0.5)
```

So, if you were to flip a fair coin ten times, there is an approximately 20.50781% chance that you will get tails six times.

**Note: for both bin_choose() and bin_probability(), the input for k/success can be a vector with multiple values.**

## Function bin_distribution()

This function allows users to visualize in a table the probabilities of all possible successes in a fixed number of trials. 

To continue:
```{r}
bin_distribution(trials = 10, prob = 0.5)
```

In addition to seeing this information in a data frame, it is also possible to plot it!

Taking our same table from above:
```{r}
distribution <- bin_distribution(trials = 10, prob = 0.5)
plot(distribution)
```

## Function bin_cumulative()

This function displays the same data table as bin_distribution(), except with an added column showing the cumulative probability of the given successes. 

```{r}
bin_cumulative(trials = 10, prob = 0.5)
```

To expand a little bit, the value in the cumulative column for success = 6 is representing the probability that in 10 flips, you get 0 tails, or 1 tail, or 2 tails, or 3 tails, etc -- all the way until 6. The high probability of ~82.8% makes sense.

Just like bin_distribution(), this table can also be easily plotted.
```{r}
cumulative_distribution <- bin_cumulative(trials = 5, prob = 0.5)
plot(cumulative_distribution)
```

##  Function bin_variable()

This function checks that the input values for trials and probability are valid, and returns a printed list of those values. 

```{r}
bin_variable(trials = 10, prob = 0.5)
```

There is also a linked summary function that prints the listed summary measures of the given binomial variable. 

It can be invoked like this:
```{r}
binomial_variable <- bin_variable(trials = 10, p = 0.5)
binomial_summary <- summary(binomial_variable)
binomial_summary
```


## Functions of Measures 

**This is a group of several functions that compute various summary measures of the binomial distribution.**

bin_mean() calculates the mean of the binomial distribution.
```{r}
bin_mean(trials = 10, p = 0.5)
```

bin_variance() calculates the variance of the binomial distribution.
```{r}
bin_variance(10, 0.5)
```



bin_mode() calculates the mode of the binomial distribution.
```{r}
bin_mode(10, 0.5)
```

12

bin_skewness() calculates the skewness of the binomial distribution.
```{r}
bin_skewness(10, 0.5)
bin_skewness(10, 0.3)
bin_skewness(10, 0.8)
```


bin_kurtosis() calculates the kurtosis of the binomial distribution.
```{r}
bin_kurtosis(10, 0.5)
```


