---
title: "Workout01"
author: "Lucy Brooke"
date: "3/10/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, echo=FALSE, results='hide'}
library('readr')
library('ggplot2')
library('dplyr')
```

```{r, echo=FALSE}
shots_data <- read.csv("/Users/lucy/Desktop/hw-stat133/workout01/data/shots-data.csv")
```

```{r overall shooting percentage table, echo=FALSE}

overall_shooting_perc <- shots_data %>%
  select(name, shot_made_flag) %>%
  group_by(name) %>%
  mutate(total = length(shot_made_flag)) %>%
  mutate(made = sum(shot_made_flag == "shot_yes")) %>%
  mutate(perc_made = (made/total)*100) %>%
  select(name, total, made, perc_made) %>%
  group_by(name, total, made, perc_made) %>%
  summarise() %>%
  arrange(desc(perc_made))

overall_shooting_perc
```

```{r 2pt shooting percentage table, echo=FALSE}
two_pt_shooting_perc <- shots_data %>%
  select(name, shot_made_flag, shot_type) %>%
  group_by(name) %>%
  mutate(total = sum(shot_type == "2PT Field Goal")) %>%
  filter(shot_type == "2PT Field Goal") %>%
  mutate(made = sum(shot_made_flag == "shot_yes")) %>%
  mutate(perc_made = (made/total)*100) %>%
  select(name, total, made, perc_made) %>%
  group_by(name, total, made, perc_made) %>%
  summarise() %>%
  arrange(desc(perc_made))


two_pt_shooting_perc
```

```{r 3pt shooting percentage table, echo=FALSE}
three_pt_shooting_perc <- shots_data %>%
  select(name, shot_made_flag, shot_type) %>%
  group_by(name) %>%
  mutate(total = sum(shot_type == "3PT Field Goal")) %>%
  filter(shot_type == "3PT Field Goal") %>%
  mutate(made = sum(shot_made_flag == "shot_yes")) %>%
  mutate(perc_made = (made/total)*100) %>%
  select(name, total, made, perc_made) %>%
  group_by(name, total, made, perc_made) %>%
  summarise() %>%
  arrange(desc(perc_made))

three_pt_shooting_perc
```



# **Introduction**

Stephen Curry's name -- even to the most casual of NBA basketball fans -- never fails to conjure up an image of a three-pointer swishing through the net. He is widely regarded as one of the greats, a [record breaker](https://www.basketball-reference.com/leaders/fg3_season.html), and according to Basketball Reference, he's a [point guard](https://www.basketball-reference.com/players/c/curryst01.html). But according to his data, does he play like a point guard?

This report primarily compares the shooting data of two players on the NBA team the Golden State Warriors: Stephen Curry and Klay Thompson. 

# **Background**
The Golden State Warriors have dominated the NBA in recent years. The season that this dataset was created (2016-17), the team won the NBA championships. In 2017-18, they placed second; in the 2018-19 season, they won again. Unsurprisingly, all the players in this report are considered to be some of the best currently playing. 

As this report involves player positions and/or expected responsibilities on the court, here is a quick breakdown. 

[Andre Iguodala](https://www.basketball-reference.com/players/i/iguodan01.html) is a **small forward**. 

[Draymond Green](https://www.basketball-reference.com/players/g/greendr01.html) is a **power forward**.

[Kevin Durant](https://www.basketball-reference.com/players/d/duranke01.html) is a **small forward/power forward/shooting guard**. For the season in question, he played as a [power forward](https://www.basketball-reference.com/teams/GSW/2017.html).

[Klay Thompson](https://www.basketball-reference.com/players/t/thompkl01.html) is a **shooting guard**. Notably, a shooting guard "is a scoring position and is often be filled by the team's best outside shooter" (https://www.basketballforcoaches.com/basketball-positions/), although a great shooting guard should be able to make shots from anywhere on the court.

[Stephen Curry](https://www.basketball-reference.com/players/c/curryst01.html) is a **point guard**. A traditional point guard is responsible for distributing the ball to those in scoring positions, like a shooting guard. However, a point guard should also have a good outside/3pt shot. (https://www.basketballforcoaches.com/basketball-positions/)

# **Data/Analysis**
Here are shot charts for all five of the players considered from the Golden State Warriors. 

![Shot charts for all five players.](/Users/lucy/Desktop/hw-stat133/workout01/images/gsw-shot-charts.pdf)

When taking a closer look, Thompson and Curry's shot charts seem practically interchangeable. Curry understandably sees a higher concentration of shots outside the three-point line; but other than that, they seem to be filling the same role.

![Shot data for Klay Thompson, a shooting guard.](/Users/lucy/Desktop/hw-stat133/workout01/images/klay-thompson-shot-chart.pdf)
![Shot data for Stephen Curry, a point guard.](/Users/lucy/Desktop/hw-stat133/workout01/images/stephen-curry-shot-chart.pdf)
```{r}
overall_shooting_perc
```

For overall shooting percentages, Thompson and Curry are -- surprisingly or unsurprisingly -- not at the top of the list. They do have, however, very similar values. Klay Thompson is squarely in the middle with roughly a 47.1% success rate, and Stephen Curry is right below him with a 46.7%. 
 
Perhaps even more interestingly, both players took almost exactly the same number of total shots; Curry ended the season with a whopping 1250, Thompson with 1220.

The fact that Curry, who as a point guard is supposed to be distributing/creating scoring chances for other players/moving the ball down the court, is taking essentially the same number of scoring chances as a shooting guard suggests that Curry is not playing a traditional point guard role. 
```{r}
two_pt_shooting_perc %>%
  arrange(desc(total))
```
Here, the total 2pt field goals taken by Curry represent a more standard breakdown of points -- Durant and Thompson take the most 2pt shots, while Curry stays more outside.
```{r}
three_pt_shooting_perc
```

When the data is limited to three point shots, the similarity of the roles Thompson and Curry are playing continues to stand out. Ideally both a traditional point guard and a shooting guard should have a strong outside shot, so the similarity here should not be too unusual. Funnily enough, even though Curry has a reputation for excelling at 3-pointers, Thompson beats him for accuracy (but Curry still has him -- and everybody -- for total number of threes).


# **Conclusion**
From the shot charts and breakdowns of shooting data, it seems that Stephen Curry is eschewing a traditional point guard role in favor of a style of play that mixes elements of point guard and shooting guard.


